---
title: "ars_cold_model"
author: "Ranae Dietzel"
date: "September 19, 2017"
output: pdf_document
---  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(lubridate)
library(modelr)
library(stringr)
library(aqp)
library(soilDB)
library(ggmap)
library(maps)
library(mapdata)
library(devtools)
library(readr)
```  

```{r map of data, echo=FALSE, warning=FALSE, message=FALSE}  
ars <- read_csv("../data/ars/ars_latlong.csv")%>%
  select(`Experimental Unit`, `Latitude of weather station, decimal degrees`, 
         `Longitude of weather station, decimal degrees`)%>%
  rename(name = `Experimental Unit`, lon = `Longitude of weather station, decimal degrees`, lat = `Latitude of weather station, decimal degrees`)

ars_box <- make_bbox(lon = ars$lon, lat = ars$lat, f = .1)

sq_map <- get_map(location = ars_box, maptype = "toner", source = "stamen")

ggmap(sq_map) + 
  geom_point(data = ars, mapping = aes(x = lon, y = lat), color = "red", size = 3.5)+
geom_point(data = ars, mapping = aes(x = lon, y = lat), color = "black", size = 3.0)  
```
1. Read and select ARS flux data  

```{r Read and select ARS flux data}
ars_fluxes<-read_csv("../data/ars/ars_fluxes.csv", skip = 2)

fluxes<-ars_fluxes%>%
  select(`Experimental Unit ID`, Date, N2O, Air_Temp, Soil_Temp, Soil_Moisture)%>%
  mutate(Date = mdy_hms(Date))%>%
  separate(`Experimental Unit ID`, c("state","exp"), sep = 3)
```  

2. Read and select ARS weather data  

```{r, read and select ars weather data}
ars_weather<-read_csv("../data/ars/ars_weather.csv", skip = 2)

weather<-ars_weather%>%
  select(`Site ID`, Date, `Temperate, air, daily, maximum, degrees Celsius`, `Temperate, air, daily, minimum, degrees Celsius`, Precip, `Total_Net_Radn`)%>%
  mutate(Date = mdy_hms(Date))%>%
  separate(`Site ID`, c("state","exp"), sep = 3)%>%
  filter(state != "ORP")%>%
  select(-exp)%>%
  mutate(Total_Net_Radn = as.numeric(Total_Net_Radn))
```  

3. Read and select ARS longitude and latitude  

```{r, read and select ars weather station data for longitude and latitude}
ars_latlong<-read_csv("../data/ars/ars_latlong.csv")

latlong<-ars_latlong%>%
  select(`Experimental Unit`, `Latitude of weather station, decimal degrees`, `Longitude of weather station, decimal degrees`)%>%
  rename(site = `Experimental Unit`, lat = `Latitude of weather station, decimal degrees`, lon = `Longitude of weather station, decimal degrees`)%>%
  #separate(`Experimental Unit`, c("state","exp"), sep = 3)%>%
  #filter(state != "ORP")%>%
  distinct(lat, .keep_all = TRUE)%>%
  mutate(lat = round(lat, digits = 4), lon = round(lon, digits = 4))


  

write.csv(latlong, file = "fordaymet.csv", row.names = FALSE, quote = FALSE, col.names = FALSE)
```  

4. Read and select DAYMET data for all the cold sites (Mandan, Morris, Roseville, University Park, West Lafayette)

```{r, read and select daymet ars data}		
daymet_mandan<-read_csv("../data/ars/daymet_mandan.csv", skip = 7)		
daymet_mandan$town<-"Mandan"		
daymet_morris<-read_csv("../data/ars/daymet_morris.csv", skip = 7)		
daymet_morris$town<-"Morris"		
daymet_roseville<-read_csv("../data/ars/daymet_roseville.csv", skip = 7)		
daymet_roseville$town<-"Roseville"		
daymet_university_park<-read_csv("../data/ars/daymet_university_park.csv", skip = 7)		
daymet_university_park$town<-"University_Park"		
daymet_west_lafayette<-read_csv("../data/ars/daymet_west_lafayette.csv", skip = 7)		
daymet_west_lafayette$town<-"West_Lafayette"		
		
daymet<-rbind(daymet_mandan, daymet_morris, daymet_roseville, daymet_university_park, daymet_west_lafayette)		
		
daymet$date<-as.Date(strptime(paste(daymet$year, daymet$yday), format="%Y%j"))		
colnames(daymet)<-c("year", "yday", "daymet_prcp", "daymet_radn", "daymet_tmax", "daymet_tmin", "town", "date")	
```  

```{r, read and download DAYMET data}

daymet_all<-download_daymet_batch(file_location = 'fordaymet.csv',
                                  start = 2003, 
                                  end = 2011, 
                                  internal = TRUE)

```

5. Read and select soils series for each site  

```{r, read and select soils series for experiments}
ars_soilseries<-read_csv("../data/ars/ars_soilseries.csv")

soils<-ars_soilseries%>%
  select(`Experimental Unit`, `Soil series`)%>%
  rename(series = `Soil series`)%>%
  separate(`Experimental Unit`, c("state","exp"), sep = 3)%>%
  filter(state != "ORP")%>%
  #select(-exp)%>%
  distinct()%>%
  mutate(series = word(series, 1))%>%
  na.omit()
```  

6. Get soil sand silt, clay, organic C, and pH from SSURGO  

```{r, get and add soil data}

our_soils<-soils%>%
  distinct(series)%>%
  mutate(dom_soil = word(series, 1))%>%
  distinct(dom_soil)%>%
  na.omit()

get_soil_data <- function(series){
  horizons(fetchKSSL(series))%>%
    select(pedon_key, hzn_desgn, sand, silt, clay, oc, ph_h2o)%>%
    filter(grepl('A', hzn_desgn))%>%
    select(-c(hzn_desgn, pedon_key))%>%
    summarise_each(funs(mean(., na.rm = TRUE)))%>%
    mutate(series = series)  
}  

brute_errors<-lapply(our_soils$dom_soil, function(series) try(get_soil_data(series)))

ars_soils<-bind_rows(Filter(function(series) !inherits(series, "try-error"), brute_errors))
```  

7. Join all the data together  

```{r, join and filter fluxes, weather, and location data}
ars<-left_join(weather, fluxes, by=c("state", "Date"))
ars<-left_join(ars, soils, by = c("state", "exp"))
ars<-left_join(ars, ars_soils, by = "series" )
ars<-left_join(ars, latlong, by = "state")


colnames(ars)<-c("site", "date", "max_temp", "min_temp", 
                 "precip", "radn", "exp", "N2O", "air_temp",
                 "soil_temp", "soil_moisture", "series", 
                 "sand", "silt", "clay", "oc", "ph_h2o", "exp.y", "lat", "long")

ars<-ars%>%
  select(-exp.y)%>%
  mutate(year = year(date), month = month(date), day = yday(date))
```  

8. Choose only the sites that freeze  

```{r, looking only at sites that freeze}
ars_cold<-ars %>% filter(site %in% c("INA", "INT", "INW", "MNM", "MNR", "NDM", "NEM", "NVN", "PAH"))%>%
  mutate(town = ifelse((site %in% c("INA", "INT", "INW")), "West_Lafayette",
                       ifelse((site == "MNM"), "Morris",
                              ifelse((site == "MNR"), "Roseville",
                                     ifelse((site %in% c("NDM", "NEM", "NVN")), "Mandan",
                                            ifelse((site == "PAH"), "University_Park", "NA" ))))), 
                                            date = as.Date(date))

ars_cold<-ars_cold%>%
  mutate(spring_year = ifelse((date >"2003-06-02" & date < "2004-06-02"), 2004,
                        ifelse((date >"2004-06-02" & date < "2005-06-02"), 2005,
                               ifelse((date >"2005-06-02" & date < "2006-06-02"), 2006,
                                      ifelse((date >"2006-06-02" & date < "2007-06-02"), 2007,
                        ifelse((date >"2007-06-02" & date < "2008-06-02"), 2008,
                          ifelse((date >"2008-06-02" & date < "2009-06-02"), 2009,
                             ifelse((date >"2009-06-02" & date < "2010-06-02"), 2010,
                                ifelse((date >"2010-06-02" & date < "2011-06-02"), 2011, 0)))))))))

day<-c(155:365, 1:154)
new_days<-1:365

new_year<-data.frame(day, new_days)

ars_cold<-left_join(ars_cold, new_year, by = "day")  

ggplot(ars_cold, aes(x=town, y=N2O))+
  geom_jitter(alpha=.3)
```

###Make annual-based variables

```{r, Make average spring N2O for a response variable}

#may want to try converting average N2O into cumulative N2O like Wagner-Riddle
ars_spring<-ars_cold%>% 
  filter(day<150)%>%
  select(-exp, -series)%>%
  group_by(site,town, year)%>%
  summarise(avg_N2O = (mean(N2O, na.rm = TRUE)))  

ggplot(ars_spring, aes(x=year, y=avg_N2O))+
  geom_point()+
  facet_wrap(~town)+
  ggtitle("Avg spring N2O (Jan-May)")
```

```{r, Daymet-based cold factor}
how_cold<-daymet%>%
  mutate(spring_year = ifelse((date >"2003-06-02" & date < "2004-06-02"), 2004,
                              ifelse((date >"2004-06-02" & date < "2005-06-02"), 2005,
                                     ifelse((date >"2005-06-02" & date < "2006-06-02"), 2006,
                                            ifelse((date >"2006-06-02" & date < "2007-06-02"), 2007,
                                                ifelse((date >"2007-06-02" & date < "2008-06-02"), 2008,
                                                ifelse((date >"2008-06-02" & date < "2009-06-02"), 2009,
                                                ifelse((date >"2009-06-02" & date < "2010-06-02"), 2010,
                                  ifelse((date >"2010-06-02" & date < "2011-06-02"), 2011, 0)))))))))%>%
  filter(daymet_tmin < 0)%>%
  group_by(town, spring_year)%>%
  summarise(cold_sum = sum(daymet_tmin))%>%
  filter(spring_year != 0)%>%
  rename(year = spring_year)  

ggplot(how_cold, aes(x=year, y=cold_sum, color = town))+
  geom_point(size=2)+
  ggtitle("Sum of temperatures below zero (November-May)")
```  

```{r, Daymet-based warm factor}
how_warm<-daymet%>% 
  filter(yday<150)%>%
  group_by(town, year)%>%
  summarise(sum_tmax = sum(daymet_tmax))  

ggplot(how_warm, aes(x=year, y=sum_tmax, color=town))+
  geom_point(size=2)+
  ggtitle("Sum of max temps (Jan-May)")
```

```{r, Bring annual variables together}
annual<-ars_spring%>%
  left_join(how_cold, by = c("town", "year"))%>%
  left_join(how_warm, by = c("town", "year"))

ggplot(annual, aes(x=sum_tmax, y=avg_N2O))+
  geom_point()+
  geom_smooth(method = "lm")  

ggplot(annual, aes(x=cold_sum, y=avg_N2O))+
  geom_point()+
  geom_smooth(method = "lm")
```  

```{r, Add soil characteristics back in}
the_data<-ars_cold%>%
  select(site, sand, silt, clay, oc, ph_h2o, lat)%>%
  group_by(sand)%>%
  distinct(.keep_all=TRUE)%>%
  group_by(site)%>%
  summarize_each(funs(mean(., na.rm = TRUE)))%>%
  full_join(annual, by = "site")%>%
  mutate(temp_dif = sum_tmax - cold_sum)
```

```{r, how bout some relationships}
pairs(the_data[, c(2:6, 10:12)])
```

```{r, how bout some models}

library(MASS)

fit<-lm(avg_N2O ~ cold_sum + sum_tmax + oc + clay + ph_h2o, data=the_data)
step<- stepAIC(fit, direction="both")

step$anova

fit2<-lm(avg_N2O ~ cold_sum + sum_tmax + oc, data=the_data)


library(boot)

glm.fit = glm(avg_N2O ~ cold_sum + sum_tmax + oc, data=the_data)

cv.glm(the_data, glm.fit)$delta

nonan<-as.data.frame(unclass(na.omit(the_data)))

library(caret)
train_control<-trainControl(method="LOOCV")

model1<-train(avg_N2O ~ sum_tmax + cold_sum + oc + clay, data=nonan, trControl=train_control, method="lm")

model2<-train(avg_N2O ~ sum_tmax + oc + clay + oc*clay, data=nonan, trControl=train_control, method="lm")

model3<-train(avg_N2O ~ sum_tmax + oc*sum_tmax, data=nonan, trControl=train_control, method="lm")
```




