---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(lubridate)
```

```{r read and select ars flux data}
ars_fluxes<-read_csv("../data/ars_fluxes.csv", skip = 2)

fluxes<-ars_fluxes%>%
  select(`Experimental Unit ID`, Date, N2O, Air_Temp, Soil_Temp, Soil_Moisture)%>%
  mutate(Date = mdy_hms(Date))%>%
  separate(`Experimental Unit ID`, c("state","exp"), sep = 3)

head(fluxes)
```

```{r read and select ars weather data}
ars_weather<-read_csv("../data/ars_weather.csv", skip = 2)

weather<-ars_weather%>%
  select(`Site ID`, Date, `Temperate, air, daily, maximum, degrees Celsius`, `Temperate, air, daily, minimum, degrees Celsius`, Precip, `Total_Net_Radn`)%>%
  mutate(Date = mdy_hms(Date))%>%
  separate(`Site ID`, c("state","exp"), sep = 3)%>%
  filter(state != "ORP")

head(weather)
```
```{r read and select daymet ars data, warning=TRUE}
daymet_mandan<-read_csv("../data/daymet_mandan.csv", skip = 7)
daymet_mandan$town<-"Mandan"
daymet_morris<-read_csv("../data/daymet_morris.csv", skip = 7)
daymet_morris$town<-"Morris"
daymet_roseville<-read_csv("../data/daymet_roseville.csv", skip = 7)
daymet_roseville$town<-"Roseville"
daymet_university_park<-read_csv("../data/daymet_university_park.csv", skip = 7)
daymet_university_park$town<-"University_Park"
daymet_west_lafayette<-read_csv("../data/daymet_west_lafayette.csv", skip = 7)
daymet_west_lafayette$town<-"West_Lafayette"

daymet<-rbind(daymet_mandan, daymet_morris, daymet_roseville, daymet_university_park, daymet_west_lafayette)

daymet$date<-as.Date(strptime(paste(daymet$year, daymet$yday), format="%Y%j"))
colnames(daymet)<-c("year", "yday", "daymet_prcp", "daymet_radn", "daymet_tmax", "daymet_tmin", "town", "date")

#daymet<-mutate(daymet, if_else )

head(daymet)
```


```{r join and filter fluxes and weather data}
ars<-full_join(weather, fluxes, by=c("state", "Date"))

colnames(ars)<-c("site", "nada", "date", "max_temp", "min_temp", "precip", "radn", "exp", "N2O", "air_temp", "soil_temp", "soil_moisture")

head(ars)
```
```{r check these data out}

ggplot(fluxes, (aes(x = state, y = N2O)))+
  geom_jitter(alpha = .3)
```

```{r look for sites that arent cold enough}
ggplot(ars, (aes(x = date, y = max_temp)))+
  geom_point(alpha = .3)+
  geom_hline(yintercept=0)+
  facet_wrap(~site)
```
```{r remove warm sites}
ars_cold<-ars %>% filter(site %in% c("INA", "INT", "INW", "MNM", "MNR", "NDM", "NEM", "NVN", "PAH"))%>%
   mutate(town = ifelse((site %in% c("INA", "INT", "INW")), "West_Lafayette",
                        ifelse((site == "MNM"), "Morris",
                               ifelse((site == "MNR"), "Roseville",
                                      ifelse((site %in% c("NDM", "NEM", "NVN")), "Mandan",
                                            ifelse((site == "PAH"), "University_Park", "NA"                                                                                           ))))))%>%
  mutate(month=month(date, label=TRUE), year = as.factor(year(date)), date = as.Date(date))%>%
  filter(N2O < 1500 | is.na(N2O)) 

ggplot(ars_cold, (aes(x = date, y = max_temp)))+
  geom_point(alpha = .2)+
  geom_hline(yintercept=0)+
  facet_wrap(~town)
```  

```{r ars plus daymet}  
both_ars_cold<-right_join(daymet, ars_cold, by = c("date", "town"))
```

```{r cold fluxes}
ggplot(ars_cold, (aes(x = town, y = N2O)))+
  geom_jitter(alpha = .3)
```
```{r N20 over the years}
ggplot(ars_cold, (aes(x = date, y = N2O)))+
  geom_point(alpha = .3)+
  facet_wrap(~town, scales = "free")
``` 
```{r N2O by month}
ggplot(data = filter(ars_cold, month %in% c("Jan", "Feb", "Mar", "Apr", "May")) , aes(x=month, y=N2O, color = year))+
  geom_boxplot()+
  facet_wrap(~town, scales = "free", nrow=5)
```

```{r N2O and radiation in Mandan}
ggplot(data=filter(ars_cold, town == "Mandan", air_temp != "NA"), aes(x=air_temp, y=N2O, color = year))+
  geom_jitter(alpha=.3)#+
 # facet_wrap(~year, scales = "free")
head(ars_cold)
```

```{r narrow down to spring}
ars_spring<-ars_cold%>% filter(month %in% c("Jan", "Feb", "Mar", "Apr", "May"))

head(ars_spring)

```

```{r Morris spring N20}
ggplot()+
  geom_point(data = filter(both_ars_cold, town == "Morris"& year.y %in% c(2004:2008)), aes(x=yday, y=N2O))+
  geom_line(data = filter(both_ars_cold, town == "Morris"& year.y %in% c(2004:2008)),
            aes(x=yday, y=(daymet_tmax*10)), color="red")+
  facet_wrap(~year.y)
```  
```{r IN spring N20}
ggplot()+
  geom_point(data = filter(both_ars_cold, town == "West_Lafayette"& year.y %in% c(2004:2012)), aes(x=yday, y=N2O))+
  geom_line(data = filter(both_ars_cold, town == "West_Lafayette"& year.y %in% c(2004:2012)),
            aes(x=yday, y=(daymet_tmax*10)), color="red")+
  facet_wrap(~year.y)
```    

```{r Mandan spring N20}
ggplot()+
  geom_point(data = filter(both_ars_cold, town == "Mandan"& year.y %in% c(2004:2012)), aes(x=yday, y=N2O))+
  geom_line(data = filter(both_ars_cold, town == "Mandan"& year.y %in% c(2004:2012)),
            aes(x=yday, y=(daymet_tmax)), color="red")+
  facet_wrap(~year.y)
```  
##What's important  

What is the definition of a peak?  
  Number of days over average for each year
  Greatest day over average for each year  
  Greatest single value  
  
What is the definition of a thaw?
  Soil temp above 0C - for how long?  
  If no soil temp, will need to choose air temp  
  Set day after which "thaw" is not an issue  

What are antecedant conditions?  
  Lowest temp 
  Freezing hours  
  Duration of freeze
  Frequency of freeze-thaws  
  
What are my assumptions?
  Soil is saturated at thaw 
  Thaw speed is the same? 
  
How does radiation play into everything?  
Temp on day of peak?  

What do I do about missed peaks?  

Are IN peaks from fertilizer at planting? Probably.

I need more data from Canada.
My NY data?
Iowa data?
Duxbury data?
NEED MORE DATA.

Look at Johnson's paper.

##Other questions  

Latitude and N2O
Soil texture  
Soil CN  

```{r kbs data}  
datain<-read_csv("../data/KBS.csv", skip=19)
colnames(datain)<-c("date", "trt", "rep", "CH4", "CO2", "N2O_N", "year", "temp", "crop")
datain$day<-yday(datain$date)
```

```{r kbs raw fluxes in N grams per ha per day}
ggplot(datain, aes(x=day, y = N2O_N))+
         geom_jitter()+
         facet_wrap(~year)
```
  
  
http://www.n2o.net.au/information/for-researchers/ rDataOne  

Flechard C.R., P. Ambus, U. Skiba, R.M. Rees, A. Hensen, A. van Amstel, A. van den Pol-van Dasselaar, J.-F. Soussana, M. Jones, J. Clifton-Brown, A. Raschi, L. Horvath, A. Neftel, M. Jocher, C. Ammann, J. Leifeld, J. Fuhrer, P. Calanca, E. Thalman, K. Pilegaard, et al. (2007) Effects of climate and management intensity on nitrous oxide emissions in grassland systems across Europe. Agriculture, Ecosystems and Environment, 121(1-2) 135-152  

Pihlatie M, 2007: Nitrous oxide emissions from selected natural and managed northern ecosystems. University of Helsinki, Department of Environmental Sciences. PhD thesis  

Estimates neglecting cold regions + thaw periods miss 17 to 28% of the story. 
  
